---
title: "Survival Probability of Titanic Passengers"
author: "Charmeine Ko"
date: "7/18/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Overview
The project aims to model the survival probability of a Titanic passenger based on the "Titanic" data set from R package "datasets". "Titanic" provides information on the fate of passengers on the famous tragedy of "Titanic" ocean liner, summarized by class status, sex, age, and survival. 
  
The Shiny application is available [here](https://lko26.shinyapps.io/DataProducts/). 
  
Source codes are availabel on [Github](https://github.com/shumengk/DataProducts).

## For Users
There are three input options available on the application:  
   
1. Class (default: 1st)  
2. Sex (default: Male)  
3. Age (default: Child)  
  
Upon selecting desired features, the user will press the "Submit" button to obtain the survival probability of a passenger with the chosen attributes. In addition, survival probability distribution is plotted for each colored by predictor value, with the chosen feature highlighted.  

## Titanic Passenger Data
The original data set is a cross-tabulation table, which aggregates individuals that have the specific characteristics and records the frequency.

```{r libraries, echo=FALSE, message=FALSE, cache=TRUE}
library(datasets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
```
```{r preprocess, echo=TRUE, message=FALSE}
str(data.frame(Titanic))
```
## Data Processing
For modeling purpose, the data set is first transformed into the long format, meaning that each observation occupies one row.
```{r transform, echo=TRUE, message=FALSE}
Titanic1 <- data.frame(Titanic) %>%
    rowwise() %>%
    mutate(Freq = list(seq(1, Freq))) %>%
    ungroup() %>%
    unnest(cols=c(Freq)) %>%
    select(-Freq)
str(Titanic1)
```
## Sample Data Analysis
Logistic regression is used to model the survival probability of Titanic passengers. The following is the survival probability of a 1st Class, female, child passenger:  
```{r modeling, echo = FALSE, cache=TRUE}
logit2prob <- function(logit){
    odds <- exp(logit)
    prob <- odds / (1 + odds)
    return(prob)
}
mod <- glm(Survived~Class+Sex+Age, data=Titanic1, family="binomial")
pred <- predict(mod, Titanic1)
Titanic1$Pred <- logit2prob(pred)

```
```{r prediction, echo = TRUE}
# mod is the logistic regression model, and logit2prob converts logit output to probability
logit2prob(predict(mod, data.frame(Class="1st", Sex="Female", Age="Child")))
```
The sample probability distribution plots are shown below, note that the curves for "1st Class", "Female", and "Child" are highlighted, in comparison to the other values.
```{r plot_prep, echo = FALSE, cache = TRUE, message = FALSE}
p1 <- ggplot(Titanic1, aes(x=Pred, fill=Class, alpha = Class == "1st")) +
            geom_density() + theme_bw() +
            scale_alpha_manual(values=c(0.5, 1), guide=FALSE) + 
            labs(title="Class") +
            xlab("Probability")
p2 <- ggplot(Titanic1, aes(x=Pred, fill=Sex, alpha = Sex == "Female")) +
            geom_density() + theme_bw() +
            scale_alpha_manual(values=c(0.5, 1), guide=FALSE) + 
            labs(title="Gender") +
            xlab("Probability")
p3 <-  ggplot(Titanic1, aes(x=Pred, fill=Age, alpha = Age == "Child")) +
            geom_density() + theme_bw() +
            scale_alpha_manual(values=c(0.5, 1), guide=FALSE) + 
            labs(title="Age Group") +
            xlab("Probability")
ggarrange(p1, p2, p3)
```
